---
title: "Survivorship of swimming larvae + settlement"
author: "Ana M. Palacio & Xaymara Serrano"
date: "`r Sys.Date()`"
output:
  html_document:
      toc: true
      df_print: paged
      theme: united
      code_folding: hide
bibliography: packages.bib
nocite: '@*'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

# General project set-up 

```{r libraries, results="hide"}

# Load all libraries and sources required to run the script
    library(tidyverse)
    library(ggthemes)
    library(survival)
    library(survminer)
    library(gtsummary)
    library(ggplot2)

  library(lme4)
  library(multcomp)
  library(multcompView)
  library(emmeans)
  library(effects)
  library(lmerTest)
  library(modelsummary)
  library(kableExtra)

library(sjPlot)
library(gridExtra)
library(cowplot)
   
  ggthe_bw<-theme_bw()+
    theme(panel.grid= element_blank(),
          legend.box.background = element_rect(),
          panel.background =element_rect(fill = NA, color = "black")
          )
  
  # set_theme(
  #   panel.grid= element_blank(),
  #   legend.box.background = element_rect(),
  #   #panel.background =element_rect(fill = NA, color = "black")
  # )
  
```

# 1. Survivorship 24h 

## Data

```{r}
# Data
    Survival.data1<-read.csv("Data/Survival_24h.csv", header = TRUE)
    summary(Survival.data1)
    
    Survival.data1$Replicate<-factor(Survival.data1$Replicate, ordered = F)
    
    Survival.data1$Treatment<-factor(Survival.data1$Treatment, 
                                   levels=c("Control", "Low Reef","High Reef", 
                                            "Low Port","High Port"))
    summary(Survival.data1$Treatment)
```

## Plot
```{r}
Surv<- ggplot(Survival.data1, aes (Treatment, Prop.survived)) +
  geom_boxplot ()+
  stat_summary(fun.data = "mean_cl_boot",geom = "errorbar", width = 0.2 )+
  stat_summary(fun=mean, geom="point") + 
  #geom_point(shape=21)+

  geom_jitter(alpha=0.5, shape=21)+
  theme(legend.position = "bottom")+
  scale_y_continuous(limits = c(0, 1),
                         breaks = seq(0,1,0.2),  
                         expand = c(0.01, 0.01),
                         name=("Survovorship proportion")) +
  ggthe_bw
Surv
```

## Possible GML and LMs 

```{r}
## model 1: generalized mixed model
  fit1<-glmer(cbind(Swimming, Dead) ~Treatment +(1|Replicate),
   data=Survival.data1,family="binomial")
    anova(fit1)
    summary(fit1)
    plot(fit1)
    
  odds1<-plot_model(fit1, type = "est", show.values=T, show.p = T) +theme_sjplot()
  odds1
  #ggsave(file="Outputs/odds_survival.svg", plot=odds1, width=4, height=4)
  
  effects1<-plot_model(fit1, type = "pre")  # Plots estimated marginal means (or marginal effects). 
  effects1<-effects1$Treatment + ggthe_bw +
     scale_y_continuous(limits = c(0.6, 1),
                           #breaks = seq(0,1,0.2),  
                           #expand = c(0.01, 0.01),
                           name=("Survivorship proportion"))
  effects1
  #ggsave(file="Outputs/effects_survival.svg", plot=effects1, width=4, height=4)

## model 2: generalized model
  fit2<-glm(cbind(Swimming,Dead)~Treatment, data=Survival.data1,family=binomial)
  anova(fit2)
  av.surv <- data.frame (anova(fit2))
  summary(fit2)
  
  par(mfrow=c(2,2))
  plot(fit2)
  par(mfrow=c(2,1))
  plot_model(fit2, type = "est",  show.values=T, show.p = T)
  plot_model(fit2, type = "eff")
  par(mfrow=c(1,1))
  
  anova(fit1, fit2)

## model 3: linear model
  fit3<-lm(Prop.survived~Treatment,weights=Swimming+Dead,data=Survival.data1)
  anova(fit3)
  summary(fit3)
  
  par(mfrow=c(2,2))
  plot(fit3)
  par(mfrow=c(1,1))
  
  anova(fit1, fit2, fit3, test = "Chi")

## model 4: mixed linear model
 fit4 <- lmer(Prop.survived.SQRT.ASIN.~ Treatment + (1|Replicate), 
              REML=TRUE, weights=Swimming+Dead, data= Survival.data1)
  summary(fit4)
  anova(fit4)
  ranova(fit4)
  step(fit4)
  plot(fit4)
  
# Pairwise comparisons
  # Day specific comparisons
  Sw.emmc<-emmeans(fit4, ~Treatment)
  Sw_groups<-cld(Sw.emmc)
  Sw_groups

```

## Survivorship and HZ ratio

### Data 2 

```{r}
#R1
Alive <- factor(rep(c("swimming","dead"), c(221, 29)), 
                levels=c("swimming","dead"))  
spi <- c(47, 40, 38,46, 50) # cell counts
dpi <- c(3, 10, 12, 4, 0) # cell counts
Treatment <- c("Control","High Port","High Reef","Low Port","Low Reef")
Treatment <- factor(c(rep(Treatment, spi), 
                rep(Treatment, dpi)), levels = Treatment)
dat1 <- data.frame(Alive,Treatment)
dat1$Replicate<-"R1"
summary(dat1)


#R2
Alive <- factor(rep(c("swimming","dead"), c(233, 17)), 
                levels=c("swimming","dead"))  
spi <- c(46, 47, 42,49, 49) # cell counts
dpi <- c(4, 3, 8, 1, 1) # cell counts
Treatment <- c("Control","High Port","High Reef","Low Port","Low Reef")
Treatment <- factor(c(rep(Treatment, spi), 
                rep(Treatment, dpi)), levels = Treatment)
dat2 <- data.frame(Alive,Treatment)
dat2$Replicate<-"R2"
summary(dat2)

#R3
Alive <- factor(rep(c("swimming","dead"), c(232, 18)), 
                levels=c("swimming","dead"))  
spi <- c(45, 43, 50, 44, 50) # cell counts
dpi <- c(5, 7, 0, 6, 0) # cell counts
Treatment <- c("Control","High Port","High Reef","Low Port","Low Reef")
Treatment <- factor(c(rep(Treatment, spi), 
                rep(Treatment, dpi)), levels = Treatment)
dat3 <- data.frame(Alive,Treatment)
dat3$Replicate<-"R3"
summary(dat3)

#R4
Alive <- factor(rep(c("swimming","dead"), c(245, 5)), 
                levels=c("swimming","dead"))  
spi <- c(50, 47, 50, 48, 50) # cell counts
dpi <- c(0, 3, 0, 2, 0) # cell counts
Treatment <- c("Control","High Port","High Reef","Low Port","Low Reef")
Treatment <- factor(c(rep(Treatment, spi), 
                rep(Treatment, dpi)), levels = Treatment)
dat4 <- data.frame(Alive,Treatment)
dat4$Replicate<-"R4"
summary(dat4)

#R5
Alive <- factor(rep(c("swimming","dead"), c(222, 28)), 
                levels=c("swimming","dead"))  
spi <- c(50, 40, 48, 39, 45) # cell counts
dpi <- c(0, 10, 2, 11, 5) # cell counts
Treatment <- c("Control","High Port","High Reef","Low Port","Low Reef")
Treatment <- factor(c(rep(Treatment, spi), 
                rep(Treatment, dpi)), levels = Treatment)
dat5 <- data.frame(Alive,Treatment)
dat5$Replicate<-"R5"
summary(dat5)

#R6
Alive <- factor(rep(c("swimming","dead"), c(209, 41)), 
                levels=c("swimming","dead"))  
spi <- c(43, 35, 36,47, 48) # cell counts
dpi <- c(7, 15, 14, 3, 2) # cell counts
Treatment <- c("Control","High Port","High Reef","Low Port","Low Reef")
Treatment <- factor(c(rep(Treatment, spi), 
                rep(Treatment, dpi)), levels = Treatment)
dat6 <- data.frame(Alive,Treatment)
dat6$Replicate<-"R6"
summary(dat6)

#R7
Alive <- factor(rep(c("swimming","dead"), c(219, 31)), 
                levels=c("swimming","dead"))  
spi <- c(50, 43, 39,43, 44) # cell counts
dpi <- c(0, 7, 11, 7, 6) # cell counts
Treatment <- c("Control","High Port","High Reef","Low Port","Low Reef")
Treatment <- factor(c(rep(Treatment, spi), 
                rep(Treatment, dpi)), levels = Treatment)
dat7 <- data.frame(Alive,Treatment)
dat7$Replicate<-"R7"
summary(dat7)

#R8
Alive <- factor(rep(c("swimming","dead"), c(155, 45)), 
                levels=c("swimming","dead"))  
spi <- c(41, 34, 36, 44) # cell counts
dpi <- c(9, 16, 14, 6) # cell counts
Treatment <- c("Control","High Port","High Reef","Low Port")
Treatment <- factor(c(rep(Treatment, spi), 
                rep(Treatment, dpi)), levels = Treatment)
dat8 <- data.frame(Alive,Treatment)
dat8$Replicate<-"R8"
summary(dat8)


#R9
Alive <- factor(rep(c("swimming","dead"), c(205, 45)), 
                levels=c("swimming","dead"))  
spi <- c(45, 40, 39, 41, 40) # cell counts
dpi <- c(5, 10, 11, 9, 10) # cell counts
Treatment <- c("Control","High Port","High Reef","Low Port","Low Reef")
Treatment <- factor(c(rep(Treatment, spi), 
                rep(Treatment, dpi)), levels = Treatment)
dat9 <- data.frame(Alive,Treatment)
dat9$Replicate<-"R9"
summary(dat9)

#R10
Alive <- factor(rep(c("swimming","dead"), c(125, 25)), 
                levels=c("swimming","dead"))  
spi <- c(48, 31, 46) # cell counts
dpi <- c(2, 19, 4) # cell counts
Treatment <- c("Control","High Port","High Reef")
Treatment <- factor(c(rep(Treatment, spi), 
                rep(Treatment, dpi)), levels = Treatment)
dat10 <- data.frame(Alive,Treatment)
dat10$Replicate<-"R10"
summary(dat10)

data<-rbind(dat1, dat2, dat3, dat4, dat5, dat6, dat7, dat8, dat9, dat10)
summary(data)

data$Treatment<-factor(data$Treatment, levels=c("Control", "Low Reef", 
                                            "Low Port", "High Reef","High Port"))

data$Fu.time_texp<-"1"
data$Fu.time_texp<-as.numeric(data$Fu.time_texp)

data$Fu.stat_exp<-"0"
data$Fu.stat_exp[data$Alive=="dead"]<-"1"
data$Fu.stat_exp<-as.numeric(data$Fu.stat_exp)
summary(data)
```

### Model 2: Survivorship analysis

```{r}
## Add survival object (Fit survival data using the Kaplan-Meier method)
  surv_object <- Surv(time = data$Fu.time_texp, event = data$Fu.stat_exp)
  #surv_object 

```

### Treatment model

```{r}
# Only treatment model

    # Kaplan-Meier estimator. The "log-log" confidence interval is preferred.
    survfit1 <- survfit(surv_object ~ Treatment, data = data)
    summary(survfit1)
    summary(survfit1)$table
    
    sd1<-survdiff(surv_object~Treatment, data = data)
    1 - pchisq(sd1$chisq, length(sd1$n) - 1)# pvalue
    
    
    results<-summary(survfit1, times = c(1))
    save.df <- as.data.frame(results[c("strata", "time", "n.risk", "n.event", "surv", "std.err")])
    write.csv(save.df, file = "Outputs/survival.csv")
    
   # Treatment_Only<-ggsurvplot(fit1, data = data, pval = TRUE, 
   #         conf.int = T, risk.table=T, 
   #         risk.table.y.text = FALSE,
   #         risk.table.title="Number of fragments at risk") + ggtitle("Treatment")
   # Treatment_Only
```

### Cox hazards 

```{r}
  coxfit <- coxph(surv_object ~ Treatment, data = data)
    summary(coxfit)
    coxfit
    #ggadjustedcurves(coxfit, data=Survival.data, variable = "Treatment")
    
    coxfit %>% 
      gtsummary::tbl_regression(exp = TRUE) 
  
  HazardRatio<-ggforest(coxfit, data = data)
  HazardRatio

#ggsave("Outputs/HazardRatio.svg", HazardRatio, width=5, height=4,dpi = 300)

```

* Hazard ratios. The exponentiated coefficients (exp(coef)), also known as hazard ratios, give the effect size of covariates. Confidence intervals of the hazard ratios. The summary output also gives upper and lower 95% confidence intervals for the hazard ratio (exp(coef))

* Global statistical significance of the model. p-values for three alternative tests for overall significance of the model: The likelihood-ratio test, Wald test, and score logrank statistics. These three methods are asymptotically equivalent. For large enough N, they will give similar results. For small N, they may differ somewhat. The Likelihood ratio test has better behavior for small sample sizes, so it is generally preferred.


# 2. Settlement

## Data

```{r}
# Data
    Survival.data2<-read.csv("Data/Survival_Settlement_week.csv", header = TRUE)
    summary(Survival.data2)
    
    Survival.data2$Replicate<-factor(Survival.data2$Replicate, ordered = F)
    
    Survival.data2$Treatment<-factor(Survival.data2$Treatment, 
                          levels=c("Control", "Low Reef", "High Reef", "Low Port", "High Port"))
    summary(Survival.data2$Treatment)
    
    # Long format
    Sur_long<-read.csv("Data/Survival_Settlement_week_long.csv", header = TRUE)
    summary(Sur_long)
    Sur_long$Treatment<-factor(Sur_long$Treatment, 

                                   levels=c("Control", "Low Reef","High Reef", 
                                            "Low Port","High Port"))
    Sur_long$Category<-factor(Sur_long$Category, 
                                   levels=c("Dead", "Un_settled","Settled"))
```

## Basic plots

### Plot 1: Xaymara settlememnt proportion (asumme all 15 larvae could settle)

```{r}
Settle<- ggplot(Survival.data2, aes (Treatment, Prop.settled)) +
  geom_boxplot ()+
  stat_summary(fun.data = "mean_cl_boot",geom = "errorbar", width = 0.2 )+
  stat_summary(fun=mean, geom="point") + 
  #geom_point(shape=21)+

  geom_jitter(alpha=0.5, shape=21)+
  theme(legend.position = "bottom")+
  scale_y_continuous(limits = c(0, 0.5),
                         breaks = seq(0,1,0.2),  
                         expand = c(0.01, 0.01),
                         name=("Settlement proportion")) +
  ggthe_bw
Settle

```

### Plot 2: Settlememnt proportion from alive larvae
```{r}
Settle2<- ggplot(Survival.data2, aes (Treatment, Prop.settled_alive)) +
  geom_boxplot ()+
  stat_summary(fun.data = "mean_cl_boot",geom = "errorbar", width = 0.2 )+
  stat_summary(fun=mean, geom="point") + 
  #geom_point(shape=21)+

  geom_jitter(alpha=0.5, shape=21)+
  theme(legend.position = "bottom")+
  scale_y_continuous(limits = c(0, 0.8),
                         breaks = seq(0,1,0.2),  
                         expand = c(0.01, 0.01),
                         name=("Settlement proportion")) +
  ggthe_bw
Settle2
```

### Plot 3: 1-week survivorship
```{r}
Week_surv<- ggplot(Survival.data2, aes (Treatment, Prop_survived)) +
  geom_boxplot ()+
  stat_summary(fun.data = "mean_cl_boot",geom = "errorbar", width = 0.2 )+
  stat_summary(fun=mean, geom="point") + 
  #geom_point(shape=21)+

  geom_jitter(alpha=0.5, shape=21)+
  theme(legend.position = "bottom")+
  scale_y_continuous(limits = c(0, 1),
                         breaks = seq(0,1,0.2),  
                         expand = c(0.01, 0.01),
                         name=("Survivorship proportion")) +
  ggthe_bw
Week_surv
```

### Plot 4: 1-week all stacked
```{r}
# Grouped
Spread_<-ggplot(Sur_long, aes(fill=Category, y=Number, x=Treatment)) + 
    geom_bar(position="dodge", stat="identity")

# Grouped
Stack<-ggplot(Sur_long, aes(fill=Category, y=Number, x=Treatment)) + 
    geom_bar(position="fill", stat="identity") +  ggthe_bw
Stack
```

## Possible GML and LMs for Xaymara's settlement

```{r}
## model 1: generalized mixed model
  fit5<-glmer(cbind(N_settled, 15-N_settled) ~Treatment +(1|Replicate),
   data=Survival.data2,family="binomial")
  anova(fit5)
  summary(fit5)
  plot(fit5)
  
  odds2<-plot_model(fit5, type = "est", show.values=T, show.p = T) +theme_sjplot()
  odds2
  #ggsave(file="Outputs/odds_settlememnt.svg", plot=odds2, width=4, height=4)
  
  
## model 2: generalized model
  fit6<-glm(cbind(N_settled, 15-N_settled)~Treatment,
            data=Survival.data2,family=binomial)
  anova(fit6)
  av_sett <- data.frame(anova(fit6))
  summary(fit6)
  #confint(fit6)
  par(mfrow=c(2,2))
  plot(fit6)
  par(mfrow=c(1,1))
  plot_model(fit6, type = "est", show.values=T, show.p = T)
  plot_model(fit6, type = "eff") 

anova(fit5, fit6, test = "Chi")
  
  effects2<-plot_model(fit6, type = "eff",
                       show.values=T,
                       show.p=T)
  effects2<-effects2$Treatment + ggthe_bw +
     scale_y_continuous(limits = c(0, 0.35),
                           #breaks = seq(0,1,0.2),  
                           #expand = c(0.01, 0.01),
                           name=("Settlement proportion"))
  effects2
  

## model 3: linear model
  fit7<-lm(Prop.settled ~Treatment,data=Survival.data2)
  anova(fit7)
  summary(fit7)
  par(mfrow=c(2,2))
  plot(fit7)
  par(mfrow=c(1,1))

## model 4: mixed linear model
 fit8 <- lmer(Prop.settled.SQRT.ASIN.~ Treatment + (1|Replicate), 
              REML=TRUE, data= Survival.data2)
  summary(fit8)
  anova(fit8)
  ranova(fit8)
  plot(fit8)
  

# Pairwise comparisons
  # Day specific comparisons
  Sw.emmc<-emmeans(fit8, ~Treatment)
  Sw_groups<-cld(Sw.emmc)
  Sw_groups

anova(fit5, fit6)
#anova(fit7, fit8)
```


```{r, model summary - survivrship }
Larvae_models<-list("Survivorship" = fit1, 
                   "Settlement" = fit6)
modelsummary(Larvae_models, stars = TRUE, 
             #statistic = c('std.error', 'p.value', 'conf.int'),
             title = 'Larvae performance models outputs')
#modelsummary(Larvae_models, estimate = "p.value")


# Anova models into one dataframe/table
    Larva_models <- data.frame(cbind
                    (c("NULL", "Treatment"),
                                rbind(av.surv, av_sett))) 
    colnames(Larva_models) <- c("", "DF", "Deviance",
                                 "Residual DF", "Residual Deviance")
    row.names(Larva_models) <- NULL
    
    Larva_models %>% kable("html", digits=2, 
        caption="Larvae binomial models") %>% 
      kable_styling(bootstrap_options = "striped", full_width = F) %>% 
      pack_rows(., "Survivorship (24 h)", 1, 2) %>% # groups rows with label
      pack_rows(., "Settlement (1 week)", 3, 4) 

```

!## Possible GML and LMs for alive's settlement - DO NOT USE

```{r}
## model 1: generalized mixed model
#   fit9<-glmer(cbind(N_settled, N.alive-N_settled) ~Treatment +(1|Replicate),
#    data=Survival.data2,family="binomial")
#   anova(fit9)
#   summary(fit9)
#   plot(fit9)
#   
#   plot_model(fit9, type = "est", show.values=T, show.p = T)
#   plot_model(fit9, type = "eff")
# 
# ## model 2: generalized model
#   fit10<-glm(cbind(N_settled, N.alive-N_settled)~Treatment, data=Survival.data2,family=binomial)
#   anova(fit10)
#   summary(fit10)
#   
#   par(mfrow=c(2,2))
#   plot(fit10)
#   par(mfrow=c(1,1))
#   
#   plot_model(fit10, type = "est", show.values=T, show.p = T)
#   plot_model(fit10, type = "eff")
# 
# 
# ## model 3: linear model
#   fit11<-lm(Prop.settled_alive ~Treatment,data=Survival.data2)
#   anova(fit11)
#   summary(fit11)
#   plot(fit11)

## model 4: mixed linear model
  # fit12 <- lmer(Prop.settled.SQRT.ASIN.~ Treatment + (1|Replicate), 
  #               REML=TRUE, data= Survival.data2)
  # summary(fit12)
  #   anova(fit12)
  #   ranova(fit12)
  #   plot(fit12)
    
  # Pairwise comparisons
    # Day specific comparisons
    # Sw.emmc<-emmeans(fit12, ~Treatment)
    # Sw_groups<-cld(Sw.emmc)
    # Sw_groups

```

## Possible GML and LMs for survivorship

```{r}
## model 1: generalized mixed model
  fit13<-glmer(cbind(N.alive, N_Dead) ~Treatment +(1|Replicate),
   data=Survival.data2,family="binomial")
  anova(fit13)
  summary(fit13)
  plot(fit13)

  # plot fixed effects correlation matrix
  plot_model(fit13, type = "est", show.values=T, show.p = T)
  plot_model(fit13, type = "eff")

## model 2: generalized model
  fit14<-glm(cbind(N.alive, N_Dead)~Treatment, data=Survival.data2,family=binomial)
  anova(fit14)
  summary(fit14)
  
  par(mfrow=c(2,2))
  plot(fit14)
  par(mfrow=c(1,1))
  
  plot_model(fit14, type = "est", show.values=T, show.p = T)
  plot_model(fit14, type = "eff")

## model 3: linear model
  fit15<-lm(Prop_dead ~Treatment,data=Survival.data2)
  anova(fit15)
  summary(fit15)
  par(mfrow=c(2,2))
  plot(fit15)
  par(mfrow=c(1,1))
  
  ## model 4: mixed linear model
   # fit16 <- lmer(Prop.settled.SQRT.ASIN.~ Treatment + (1|Replicate), 
   #              REML=TRUE, data= Survival.data2)
   #  summary(fit16)
   #  anova(fit16)
   #  ranova(fit16)
   #  plot(fit16)
   #  
  # Pairwise comparisons
    # Day specific comparisons
    # Sw.emmc<-emmeans(fit16, ~Treatment)
    # Sw_groups<-cld(Sw.emmc)
    # Sw_groups

```

# Figure 1

```{r}

Figure_1<- plot_grid(effects1, effects2, labels = "AUTO")
Figure_1
#ggsave(file="Outputs/Figure1.svg", plot=Figure_1, width=8, height=4)

```

# Figure S1

```{r}

Figure_S1<-plot_grid( odds1, odds2, labels = "AUTO")
Figure_S1
#ggsave(file="Outputs/FigureS1.svg", plot=Figure_S1, width=12, height=4)
```

# Packages used

```{r}
# Creates bibliography 
#knitr::write_bib(c(.packages()), "packages.bib")
```