---
title: "O. Faveolata larvae performance under sediment treatments"
author: "Ana M. Palacio & Xaymara Serrano"
date: "`r Sys.Date()`"
output:
  html_document:
      toc: true
      df_print: paged
      theme: united
      code_folding: hide
bibliography: packages.bib
nocite: '@*'
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

# General project set-up 

```{r libraries, results="hide"}

# Load all libraries and sources required to run the script
    library(tidyverse)
    library(ggthemes)
    library(survival)
    library(survminer)
    #library(gtsummary)
    library(ggplot2)

    library(lme4)
    library(multcomp)
    library(multcompView)
    library(emmeans)
    library(effects)
    library(lmerTest)
    library(modelsummary)
    library(performance)
    library(kableExtra)

    library(sjPlot)
    #library(gridExtra)
    library(cowplot)
   
  ggthe_bw<-theme_bw()+
    theme(panel.grid= element_blank(),
          legend.box.background = element_rect(),
          panel.background =element_rect(fill = NA, color = "black")
          )
```

# 1. Survivorship 24h 

## Data

```{r}
# Data
    Survival.data1<-read.csv("Data/Survival_24h.csv", header = TRUE)
    summary(Survival.data1)
    
    Survival.data1$Replicate<-factor(Survival.data1$Replicate, ordered = F)
    
    Survival.data1$Treatment<-factor(Survival.data1$Treatment, 
                                   levels=c("Control", "Low Reef","High Reef", 
                                            "Low Port","High Port"))
    summary(Survival.data1$Treatment)
```

## Possible GML and LMs 

Model 1: After comparing with other models (below) this was the best model and used in the publication. Because the model was over dispersed a dispersion parameter was added (fit1.1). 

```{r}
## model 1: generalized mixed model
  fit1<-glmer(cbind(Swimming, Dead) ~Treatment + (1|Replicate),
              data=Survival.data1,family="binomial")
    anova(fit1)
    summary(fit1)
    plot(fit1)
    check_overdispersion(fit1)
    
## unique container/observation ID for dispersion parameter 
    Survival.data1$obs <- factor(seq_len(nrow(Survival.data1)))
    Survival.data1$obs<-factor(Survival.data1$obs, ordered = F)
    
    fit1.1<-glmer(cbind(Swimming, Dead) ~Treatment + (1|obs),
                  data=Survival.data1,family="binomial")
    anova(fit1.1)
    summary(fit1.1)
    plot(fit1.1) # residuals get worse with this
    check_overdispersion(fit1.1) # but over dispersion is fixed 

## Compare binomial models with random effects
    anova(fit1, fit1.1)
    # Model 1.1 is the best AIC and is not over dispersed 
    av.surv <- data.frame (anova(fit1.1))
```

Other models tested ... they were not as good as the one chosen above 

```{r}
## model 2: generalized model - no random effects
  fit2<-glm(cbind(Swimming,Dead)~Treatment, 
            data=Survival.data1,family=binomial)
  anova(fit2)
  #av.surv <- data.frame (anova(fit2))
  summary(fit2)
  check_overdispersion(fit2) # models os overdispersed
  
  par(mfrow=c(2,2))
  plot(fit2)
  #par(mfrow=c(2,1))
  #plot_model(fit2, type = "est",  show.values=T, show.p = T)
  #plot_model(fit2, type = "eff", terms="Treatment")
  #par(mfrow=c(1,1))
  anova(fit1.1, fit2)
  #Model 1.1 is better and used for downstream multicomp and plots
```

## Pairwise comparisons and plot results

Multiple comparisons

```{r}
# Pairwise comparisons
    # To Control: 
    contrast(emmeans (fit1.1, ~ Treatment), "trt.vs.ctrl1", ref = 'Control')
    # Among all treatments
    Sw.emmc<-emmeans(fit1.1, ~Treatment)
    Sw_groups<-cld(Sw.emmc)
    Sw_groups
```
 
 Plot results
 
```{r}
# Plot results

odds1<-plot_model(fit1.1, type = "est", show.values=T, show.p = T, 
                    title = "Survivorship compared to control") +theme_sjplot()
  odds1
  #ggsave(file="Outputs/odds_survival_R1.svg", plot=odds1, width=4, height=4)
  
effects1<-plot_model(fit1.1, type = "pred", terms="Treatment", colors = c("black"), title="")  + # Plots estimated marginal means (or marginal effects). 
                      ggthe_bw +
                      scale_y_continuous(limits = c(0.6, 1),
                           #breaks = seq(0,1,0.2),  
                           #expand = c(0.01, 0.01),
                           name=("Survivorship proportion"))
  effects1
  #ggsave(file="Outputs/effects_survival.svg", plot=effects1, width=4, height=4)
```

# 2. Settlement

## Data

```{r}
# Data
    Survival.data2<-read.csv("Data/Survival_Settlement_week.csv", header = TRUE)
    summary(Survival.data2)
    
    # Random factors
    Survival.data2$Replicate<-factor(Survival.data2$Replicate, ordered = F)
    ## unique container/observation ID for dispersion parameter 
    Survival.data2$obs <- factor(seq_len(nrow(Survival.data2)))
    Survival.data2$obs<-factor(Survival.data2$obs, ordered = F)
   
    # Fixed factors
    Survival.data2$Treatment<-factor(Survival.data2$Treatment, 
                          levels=c("Control", "Low Reef", "High Reef", "Low Port", "High Port"))
    summary(Survival.data2$Treatment)
    
    # Long format
    Sur_long<-read.csv("Data/Survival_Settlement_week_long.csv", header = TRUE)
    summary(Sur_long)
    Sur_long$Treatment<-factor(Sur_long$Treatment, 

                                   levels=c("Control", "Low Reef","High Reef", 
                                            "Low Port","High Port"))
    Sur_long$Category<-factor(Sur_long$Category, 
                                   levels=c("Dead", "Un_settled","Settled"))
```

## Exploratory plots

1-week all outcomes stacked

```{r}
# Grouped
Spread_<-ggplot(Sur_long, aes(fill=Category, y=Number, x=Treatment)) + 
    geom_bar(position="dodge", stat="identity")

# Grouped
Stack<-ggplot(Sur_long, aes(fill=Category, y=Number, x=Treatment)) + 
    geom_bar(position="fill", stat="identity") +  ggthe_bw
Stack
```

## GML and for settlement

```{r}
## model 1: generalized mixed model with random effects
  fit3<-glmer(cbind(N_settled, 15-N_settled) ~Treatment + (1|Replicate),
              data=Survival.data2,family="binomial")
      check_overdispersion(fit3)
      
  fit4<-glmer(cbind(N_settled, 15-N_settled) ~Treatment + (1|obs),
              data=Survival.data2,family="binomial")
      check_overdispersion(fit4)
      
 ## model 2: generalized model     
  fit5<-glm(cbind(N_settled, 15-N_settled) ~Treatment,
              data=Survival.data2,family="binomial")
      check_overdispersion(fit5)
  
  anova(fit3, fit4, fit5, test = "Chi")
  # All models performed similarly, random effects were disccarded since glm has smaller AIC and less overdispersion (these differneces were small) 
  anova(fit5)
  summary(fit5)
  #plot(fit4)
  av_sett <- data.frame(anova(fit5))
  #confint(fit4)
  
  plot_model(fit5, type = "est", show.values=T, show.p = T)
  plot_model(fit5, type = "eff", terms="Treatment") 
```


## Pairwise comparisons and plot results

Multiple comparisons

```{r}
# Pairwise comparisons
    # To Control: 
    contrast(emmeans (fit5, ~ Treatment), "trt.vs.ctrl1", ref = 'Control')
    # Among all treatments
    Set.emmc<-emmeans(fit5, ~Treatment)
    Set_groups<-cld(Set.emmc)
    Set_groups
```
 
 Plot results
 
```{r}
# Plot results
  odds2<-plot_model(fit5, type = "est", show.values=T, show.p = T, 
                    title = "Settlement compared to control") +theme_sjplot()
  odds2
  #ggsave(file="Outputs/odds_survival_R1.svg", plot=odds1, width=4, height=4)
  
  effects2<-plot_model(fit5, type = "pred", terms="Treatment", colors = c("black"), title="") + # Plots estimated marginal means (or marginal effects). 
                      ggthe_bw +
                      scale_y_continuous(limits = c(0, 0.35),
                                        #breaks = seq(0,1,0.2),  
                                        #expand = c(0.01, 0.01),
                                        name=("Settlement proportion"))
  effects2
  #ggsave(file="Outputs/effects_survival_R1.svg", plot=effects1, width=4, height=4)
```

# Final models

```{r, model significance}

# Anova models into one data frame/table
    drop1(fit1.1,test="Chisq")
    av.surv %>% kable("html", digits=2, 
        caption="Larvae 24 survivorship") %>% 
      kable_styling(bootstrap_options = "striped", full_width = F)
    
    drop1(fit5,test="Chisq")
    av_sett %>% kable("html", digits=2, 
        caption="Settlement (1 week)") %>% 
      kable_styling(bootstrap_options = "striped", full_width = F) %>% 
      pack_rows(., "Settlement (1 week)", 1, 2) 

#To get the significance for the overall model
  #Survivorship
  #paste("Survivorship (24 h) X2 =", 1-pchisq(241.35-180.4, 46-42))

  #Settlement
  #paste("Settlement (1 week) X2 =", 1-pchisq(54.53-37.79, 43-39))
 
```

```{r, model summary - coeficients }
Larvae_models<-list("Survivorship" = fit1.1, 
                   "Settlement" = fit4)
modelsummary(Larvae_models, stars = TRUE, 
             #statistic = c('std.error', 'p.value', 'conf.int'),
             title = 'Larvae performance models outputs')
#modelsummary(Larvae_models, estimate = "p.value")
```

# Figure 1

```{r}

Figure_1<- plot_grid(effects1, effects2, labels = "AUTO")
Figure_1
#ggsave(file="Outputs/Figure1.svg", plot=Figure_1, width=8, height=4)

```

# Figure S1

```{r}

Figure_S1<-plot_grid(odds1, odds2, labels = "AUTO")
Figure_S1
#ggsave(file="Outputs/FigureS1.svg", plot=Figure_S1, width=12, height=4)
```

# 3 Respiration (non-significant)

## Data

```{r}
# Data
    data<-read.csv("Data/Respiration_larvae_v2.csv", header = TRUE)
    summary(data)
    
    data$Replicate.vial<-factor(data$Replicate.vial, ordered = F)
    data$Plate<-factor(data$Plate, ordered = F)
    data$Well<-factor(data$Well, ordered = F)
    
    data$Treatment<-factor(data$Treatment, 
                                levels=c("Control", "Low Reef","High Reef", 
                                                    "Low Port","High Port"))
    summary(data$Treatment)
```

## Basic plots - by larva

```{r}
Larvae<- ggplot(data, aes (Treatment, nmol.larva.min)) +
  #geom_boxplot ()+
  stat_summary(fun.data = "mean_cl_boot",geom = "errorbar", width = 0.2 , color="black")+
  stat_summary(fun=mean, geom="point", size=2) + 
  #geom_point(shape=21)+

  geom_jitter(alpha=0.5, shape=21, width = 0.2)+
  theme(legend.position = "bottom")+
  scale_y_continuous(#limits = c(0, 1),
                      #   breaks = seq(0,1,0.2),  
                      #   expand = c(0.01, 0.01),
                      name=expression(
                        O[2]~consumption~per~larva~(nmol~min^-1))) +
  ggthe_bw +
  facet_grid(~Timepoint)
Larvae
#ggsave(file="Outputs/S4_Oxygen.svg", plot=Larvae, width=8, height=4)

LogLarvae<- ggplot(data, aes (Treatment, log(nmol.larva.min))) +
  #geom_boxplot ()+
  stat_summary(fun.data = "mean_cl_boot",geom = "errorbar", width = 0.2 )+
  stat_summary(fun=mean, geom="point") + 
  #geom_point(shape=21)+

  geom_jitter(alpha=0.5, shape=21)+
  theme(legend.position = "bottom")+
  scale_y_continuous(#limits = c(0, 1),
                      #   breaks = seq(0,1,0.2),  
                      #   expand = c(0.01, 0.01),
                         name=("Log O2 [nmol/min/larva] corrected")) +
  ggthe_bw +
  facet_grid(~Timepoint)
LogLarvae
```

## Summary stats 

```{r}
Respiration.summary<-data %>% 
  group_by(Timepoint) %>% 
  summarize(mean = mean(nmol.larva.min, na.rm = T),
            sd = sd(nmol.larva.min, na.rm = T))
Respiration.summary

```

## LMs 

Raw data

```{r}
# Raw data model
  fit5<-lmer(nmol.larva.min ~Treatment*Timepoint + 
             (1|Replicate.vial) + (1|Plate..) + (1|Well), data=data)
      anova(fit5)
      ranova(fit5)
      #summary(fit5)
      plot(fit5)
      step(fit5)
  fit6<-lmer(nmol.larva.min ~ Timepoint + (1 | Well), data=data)
      anova(fit6)
      ranova(fit6)
      summary(fit6)
      plot(fit6)

#Pairwise comparisons
  Sw.emmc<-emmeans(fit6, ~Timepoint)
  Sw_groups<-cld(Sw.emmc)
  Sw_groups
```

log transformed data

```{r}
 fit5.1<-lmer(log(nmol.larva.min) ~Treatment*Timepoint + 
             (1|Replicate.vial) + (1|Plate..) + (1|Well), data=data)
      anova(fit5.1)
      ranova(fit5.1)
      summary(fit5.1)
      plot(fit5.1)
      step(fit5.1)
  fit6.2<-lmer(log(nmol.larva.min) ~ Timepoint + (1 | Well), data=data)
      anova(fit6.2)
      ranova(fit6.2)
      summary(fit6.2)
      plot(fit6.2)
```

Only time point and well are significant factors for respiration 

# Packages used

```{r}
# Creates bibliography 
#knitr::write_bib(c(.packages()), "packages.bib")
```